Reviewer Comments

### 2024-10-11 Manuscript.qmd
Correspondance
Starting to work on Review

format:
  pdf:
    keep-tex: true
    template-partials:
      - title.tex
    include-in-header:
      text: |
        \usepackage{academicons}
        \usepackage{fontawesome5}
        \usepackage{lineno}\linenumbers
        \usepackage[noblocks]{authblk}
        \usepackage{setspace}
        \renewcommand\Affilfont{\footnotesize}
        \newcommand{\orcid}[1]{\href{https://orcid.org/#1}{\textcolor[HTML]{A6CE39}{\aiOrcid#1}}}
        \newcommand{\mail}[1]{\href{mailto:#1}{\textcolor[HTML]{052E81}{\faEnvelope#1}}}

@@

We are thankful to the anonymous reviewers and the editors who provided valuable comments to improve the manuscript.

@@

No external funding was received.

@@

All authors contributed particular sections to the manuscript coordinated by MHKH.

Correspondence to [Maximilian H.K. Hesselbarth](mailto:mhk.hesselbarth@gmail.com)

@@

The authors declare no competing interests.

@@

This article does not contain any studies with human or animal subjects performed by any of the authors.

@@

@@

**Frazier (2019) Emerging trajectories for spatial pattern analysis in landscape ecology. Landscape Ecology 34:2073–2082.** The paper describes how interdisciplinary perspectives have contributed to spatial pattern analysis and introduces new innovative developments.

**Unnithan Kumar et al. (2022) Moving beyond landscape resistance: Considerations for the future of connectivity modelling and conservation science. Landscape Ecology 37:2465–2480.** The paper provides a historical overview of connectivity modelling with a focus on resistance surfaces and discusses current missing aspects and how to address them.

**Zuckerberg et al. (2020) A review of overlapping landscapes: Pseudoreplication or a red herring in landscape ecology? Current Landscape Ecology Reports 5:140–148.** The paper discusses theoretical and practical implications over overlapping landscapes in the context of landscape buffers of varying sizes surrounding sampling sites.

**Synes et al. (2016) Emerging opportunities for landscape ecological modelling. Current Landscape Ecology Reports 1:146–167.** The paper identifies existing gaps in landscape ecological modelling while highlighting potential emerging opportunities.

**Manel and Holderegger (2013) Ten years of landscape genetics. Trends in Ecology & Evolution 28:614–621.** Ten years after their first paper introducing landscape genetics, this paper reviews the state-of-the-art and describes main topics that have contributed to the progress of landscape genetics.

**Jung et al. (2024) An assessment of the state of conservation planning in europe. Philosophical Transactions of the Royal Society B 379:20230015.** The paper assesses where and how spatial conservation planning have been applied across Europe.

>>>>

format:
  pdf:
    template-partials:
      - title.tex
    include-in-header:
      text: |
        \usepackage{academicons}
        \usepackage{fontawesome5}
        \usepackage{lineno}\linenumbers
        \usepackage[noblocks]{authblk}
        \usepackage{setspace}
        \renewcommand\Affilfont{\footnotesize}
        \newcommand{\orcid}[1]{\href{https://orcid.org/#1}{\textcolor[HTML]{A6CE39}{\aiOrcid#1}}}
        \newcommand{\mail}[1]{\href{mailto:#1}{\textcolor[HTML]{052E81}{\faEnvelope#1}}}

@@

\noindent We are thankful to the anonymous reviewers and the editors who provided valuable comments to improve the manuscript.

@@

\noindent No external funding was received.

@@

\noindent All authors contributed particular sections to the manuscript coordinated by MHKH.
Correspondence to [mhk.hesselbarth@gmail.com](mailto:mhk.hesselbarth@gmail.com)

@@

\noindent The authors declare no competing interests.

@@

\noindent This article does not contain any studies with human or animal subjects performed by any of the authors.

@@

<!-- R1: Figure 1. The links between Spatial sampling, Spatial planning, and Genetics, towards Landscape mean? The figure would be more meaningful with an explanation of these links, either in the caption or directly on the figure. Also, for other topics (e.g. connectivity), a keyword (or short sentence) summarizing the contribution of the topic to Landscape Ecology would be great. -->

@@

\noindent _Frazier (2019) Emerging trajectories for spatial pattern analysis in landscape ecology. Landscape Ecology 34:2073–2082._ The paper describes how interdisciplinary perspectives have contributed to spatial pattern analysis and introduces new innovative developments.

\noindent _Unnithan Kumar et al. (2022) Moving beyond landscape resistance: Considerations for the future of connectivity modelling and conservation science. Landscape Ecology 37:2465–2480._ The paper provides a historical overview of connectivity modelling with a focus on resistance surfaces and discusses current missing aspects and how to address them.

\noindent _Zuckerberg et al. (2020) A review of overlapping landscapes: Pseudoreplication or a red herring in landscape ecology? Current Landscape Ecology Reports 5:140–148._ The paper discusses theoretical and practical implications over overlapping landscapes in the context of landscape buffers of varying sizes surrounding sampling sites.

\noindent _Synes et al. (2016) Emerging opportunities for landscape ecological modelling. Current Landscape Ecology Reports 1:146–167._ The paper identifies existing gaps in landscape ecological modelling while highlighting potential emerging opportunities.

\noindent _Manel and Holderegger (2013) Ten years of landscape genetics. Trends in Ecology & Evolution 28:614–621._ Ten years after their first paper introducing landscape genetics, this paper reviews the state-of-the-art and describes main topics that have contributed to the progress of landscape genetics.

\noindent _Jung et al. (2024) An assessment of the state of conservation planning in europe. Philosophical Transactions of the Royal Society B 379:20230015._ The paper assesses where and how spatial conservation planning have been applied across Europe.

### 2024-10-11 Sections/01-introduction.qmd
Fix format issues
Rename some files
Starting to work on Review

Computational science analyzes abstracted core mechanisms of research questions using data and algorithms and is one of the most important tools of modern science [@Wilson2014BestPractices].
Following, computational ecology can be defined as computational science that is used to address ecological research questions with focus on data-driven approaches and complex, adaptive systems [@Poisot2019DatabasedSynthesisdriven].
Computational ecology is crucial for landscape ecology as a research field because data is often context- and scale-dependent, making it challenging to design controllable, reproducible, and replicable experiments [@Petrovskii2012ComputationalEcology; but see @Wiersma2022ReviewLandscape for a review of landscape ecology experiments].
Additionally, because landscape ecology is a cross-disciplinary field [including e.g., social sciences, geography, and ecology and evolution\; @With2019EssentialsLandscape], the general increase of data availability [@Jaric2020IEcologyHarnessing], and the complexity of landscape systems [@Newman2019ScalingComplexity], there is the need for advanced computational methods.

Here, we introduce recent computational developments in some important topics of landscape ecology including software that implements or is potentially capable of implementing them (Fig. 1).
In this context, we are focusing on open-source software and specifically scripting languages such as R, Python, and Julia.
However, we do not provide a systematic literature review, or a general introduction to (computational) landscape ecology.
For a more comprehensive and general introduction to landscape ecology, please see @Turner2015LandscapeEcology, @Gergel2017LearningLandscape, or @With2019EssentialsLandscape.

![Overview of landscape ecology topics for which computational methods are discussed.](Figures/Fig-1.png){width=100%}

>>>>

Computational science analyzes abstracted core mechanisms of research questions using data and algorithms and is one of the most important tools of modern science [@Wilson2014BestPractices].
<!-- R1: Data-driven and model-driven -->
Following, computational ecology can be defined as computational science that is used to address ecological research questions with focus on data-driven and model-driven approaches and complex, adaptive systems [@Poisot2019DatabasedSynthesisdriven].
Computational ecology is crucial for landscape ecology as a research field because data is often context- and scale-dependent, making it challenging to design controllable, reproducible, and replicable experiments [@Petrovskii2012ComputationalEcology; but see @Wiersma2022ReviewLandscape for a review of landscape ecology experiments].
Additionally, because landscape ecology is a cross-disciplinary field [including e.g., social sciences, geography, and ecology and evolution\; @With2019EssentialsLandscape], the general increase of data availability [@Jaric2020IEcologyHarnessing], and the complexity of landscape systems [@Newman2019ScalingComplexity], there is the need for advanced computational methods.

<!-- R1: Very good idea to suggest software implementing methodological developments. It would be even better with a recapitulative table -->
Here, we introduce recent computational developments in some important topics of landscape ecology including software that implements or is potentially capable of implementing them (Fig. 1).
In this context, we are focusing on open-source software and specifically scripting languages such as R, Python, and Julia.
However, we do not provide a systematic literature review, or a general introduction to (computational) landscape ecology.
For a more comprehensive and general introduction to landscape ecology, please see @Turner2015LandscapeEcology, @Gergel2017LearningLandscape, or @With2019EssentialsLandscape.

![Overview of landscape ecology topics for which computational methods are discussed. Contains modified Copernicus Sentinel data 2024-09-07 for Sentinel data.](Figures/Fig-1.jpg)

### 2024-10-11 Sections/02-spatial-patterns.qmd
Starting to work on Review

Traditionally and at present, most of the prominent approaches that quantify spatial patterns revolve around gridded raster data using categorical LULC data on the patch mosaic model [@Frazier2017LandscapeMetrics; @Costanza2019DescribingAnalyzing].
The strength of such landscape metrics is that they are easy to apply and communicate [@Lausch2015UnderstandingQuantifying], however, limitations have also been identified. 
These limitations include shortcomings in their ability to quantify the spatial structure of landscapes, sensitivity to both spatial scale and thematic resolution, and correlation and redundancy between metrics [e.g., @Li2004UseMisuse; @Dramstad2009SpatialMetrics; @Kupfer2012LandscapeEcology].

@@

@@

Entropy measures in landscape ecology are mainly derived from information theory and thermodynamics.
They are mainly used to quantify the complexity of the landscape (spatial heterogeneity), and less often unpredictability (temporal heterogeneity), as well as scale dependence [spatio-temporal heterogeneity\; @Vranken2015ReviewUse].
However, studies show that insights gained from entropy measures depend on the formulation of the selected entropy, and on the underlining data model, e.g., the composition of categories or the co-occurrence matrix representation [@Zhao2019CalculatingSpatial].

@@

Based on concepts from thermodynamics initiated by the work of @Cushman2015ThermodynamicsLandscape and @Vranken2015ReviewUse, there has been a surge in the development of entropy-based metrics for landscape ecology in the last few years.
In the late 1800s, Boltzmann formulated a probabilistic interpretation of the second law of thermodynamics using the concepts of macrostate (the general state of a system) and microstate [the configuration of the system elements\; @Gao2019ComputationBoltzmann].
@Cushman2016CalculatingConfigurational proposed to relate the edge length to the microstate of the landscape, and use the proportion of microstates to compute the relative Boltzmann entropy of a landscape mosaic.
This approach was later generalized for calculations based on the raster surface model and point patterns [@Cushman2021EntropyLandscape].
Subsequently, @Gao2017HierarchybasedSolution proposed to use the Boltzmann entropy to quantify the complexity of a landscape surface by transforming the input raster into a series of landscape surfaces with different levels of detail (microstate) and calculating the Boltzmann entropy based on the total number of microstates that are able to generate the observed macrostate.
Recently, @Zhang2020CalculatingWasserstein extended the original definition of Boltzmann entropy to incorporate information about the adjacency of the same categories in the landscape mosaic by using the number of contiguous patches of the same category.
However, the relations between the Shannon entropy and the Boltzmann entropy and their interpretations in landscape ecology are the subject of ongoing discussion [e.g., @Stepinski2022CurbYour; @Huckeba2023MultiscaleSpatial].

@@

Surface metrics are based on the gradient surface model using gridded raster data and continuous values [@McGarigal2009SurfaceMetrics; @Cushman2010GradientParadigm]. 
These metrics were mostly adapted from microscopy and molecular physics [@McGarigal2009SurfaceMetrics; @Kedron2018SurfaceMetrics].
The gradient surface model can increase the resemblance of the data to the natural world because it allows for the inclusion of more heterogeneity within each grid cell [@Cushman2010GradientParadigm; @Kedron2018SurfaceMetrics].
Surface metrics are able to quantify various characteristics, such as roughness, skewness and kurtosis, total and relative amplitudes, curvatures of local peaks, or surface bearing area ratios [@Cushman2010GradientParadigm].
Many surface metrics have analogous landscape metrics [@Cushman2010GradientParadigm], however, using surface metrics allows for the exploration of different or additional patterns and potentially pattern-process links [@McGarigal2009SurfaceMetrics; @Borthwick2020AlternativeQuantifications].

@@

## Operations on spatial patterns

@@

Comparing spatial patterns is often used to analyze landscape dynamics, e.g., to detect changes in landscape structure over time [@Duncan2023MonitoringSpatial].
Furthermore, pattern-based search is a one-to-many operation in which the spatial signature of a query area is compared to the signatures of all other areas.
This allows for the identification of areas with similar patterns compared to the query area, e.g., areas with similar environmental conditions [@Jasiewicz2014LandscapeSimilarity].
Finally, the grouping of spatial patterns is a many-to-many operation in which spatial signatures are calculated for all areas and clustered based on their similarity [@Nowosad2021PatternbasedIdentification].

>>>>

Traditionally and at present, most of the prominent approaches that quantify spatial patterns revolve around gridded raster data using categorical LULC data on the patch mosaic model [@Frazier2017LandscapeMetrics; @Costanza2019DescribingAnalyzing].
<!-- R1: They are also naturally extracted from remote sensing products (raster data), which often makes them easier to monitor over time -->
The strength of such landscape metrics is that they are easy to apply, communicate, and can be extracted straightforwardly from raster data [e.g., remote sensing products\; @Dramstad2009SpatialMetrics; @Lausch2015UnderstandingQuantifying].
However, also limitations have been identified. 
These limitations include shortcomings in their ability to quantify the spatial structure of landscapes, sensitivity to both spatial scale and thematic resolution, and correlation and redundancy between metrics [e.g., @Li2004UseMisuse; @Dramstad2009SpatialMetrics; @Kupfer2012LandscapeEcology].

@@

<!-- R1: Missing subsection title -->
## Landscape mosaic method

@@

Entropy measures in landscape ecology are mainly derived from information theory and thermodynamics.
They are mainly used to quantify the complexity of the landscape (spatial heterogeneity), and less often unpredictability (temporal heterogeneity), as well as scale dependence [spatio-temporal heterogeneity\; @Vranken2015ReviewUse].
<!-- R1: of the selected entropy measure -->
However, studies show that insights gained from entropy measures depend on the formulation of the selected entropy measure, and on the underlining data model, e.g., the composition of categories or the co-occurrence matrix representation [@Zhao2019CalculatingSpatial].

@@

Based on concepts from thermodynamics initiated by the work of @Cushman2015ThermodynamicsLandscape and @Vranken2015ReviewUse, there has been a surge in the development of entropy-based metrics for landscape ecology in the last few years.
In the late 1800s, Boltzmann formulated a probabilistic interpretation of the second law of thermodynamics using the concepts of macrostate (the general state of a system) and microstate [the configuration of the system elements\; @Gao2019ComputationBoltzmann].
<!-- R1: the edge length -> Could you specify a bit -->
@Cushman2016CalculatingConfigurational proposed to relate the edge length to the microstate of the landscape, and use the proportion of microstates to compute the relative Boltzmann entropy of a landscape mosaic.
This approach was later generalized for calculations based on the raster surface model and point patterns [@Cushman2021EntropyLandscape].
Subsequently, @Gao2017HierarchybasedSolution proposed to use the Boltzmann entropy to quantify the complexity of a landscape surface by transforming the input raster into a series of landscape surfaces with different levels of detail (microstate) and calculating the Boltzmann entropy based on the total number of microstates that are able to generate the observed macrostate.
Recently, @Zhang2020CalculatingWasserstein extended the original definition of Boltzmann entropy to incorporate information about the adjacency of the same categories in the landscape mosaic by using the number of contiguous patches of the same category.
However, the relations between the Shannon entropy and the Boltzmann entropy and their interpretations in landscape ecology are the subject of ongoing discussion [e.g., @Stepinski2022CurbYour; @Huckeba2023MultiscaleSpatial].

@@

Surface metrics are based on the gradient surface model using gridded raster data and continuous values [@McGarigal2009SurfaceMetrics; @Cushman2010GradientParadigm]. 
These metrics were mostly adapted from microscopy and molecular physics [@McGarigal2009SurfaceMetrics; @Kedron2018SurfaceMetrics].
<!-- R1: I guess that such metrics are also very scale dependent -->
The gradient surface model can increase the resemblance of the data to the natural world because it allows for the inclusion of more heterogeneity within each grid cell [@Cushman2010GradientParadigm; @Kedron2018SurfaceMetrics].
Surface metrics are able to quantify various characteristics, such as roughness, skewness and kurtosis, total and relative amplitudes, curvatures of local peaks, or surface bearing area ratios [@Cushman2010GradientParadigm].
Many surface metrics have analogous landscape metrics [@Cushman2010GradientParadigm], however, using surface metrics allows for the exploration of different or additional patterns and potentially pattern-process links [@McGarigal2009SurfaceMetrics; @Borthwick2020AlternativeQuantifications].
Nevertheless, similar to landscape metrics, surface metrics are also scale-dependent because they are used to quantify landscape heterogeneity, which is itself scale-dependent [@Frazier2016SurfaceMetrics].

@@

## Operations on spatial patterns 
<!-- R1: I think that this could be clearer by introducing spatial patterns operation and queries as an extension of the relational data model to spatial data. I'm not sure if it is useful to use database jargon such as "one-to-many" or "many-to-many" without defining them, as it could be confusing to readers not familiar with databases. You could either avoid using these terms, define them, or illustrate the operations with a figure -->
<!-- R1: This section is missing in the summary Figure 1 -->

@@

Comparing spatial patterns is often used to analyze landscape dynamics, e.g., to detect changes in landscape structure over time [@Duncan2023MonitoringSpatial].
Furthermore, pattern-based search is a one-to-many operation in which the spatial signature of a query area is compared to the signatures of all other areas.
<!-- R1: As defined here, it looks to me that this operation is not necessarily linked to a query area, but rather to a spatial signature which can be directly specified or extracted from a query area -->
This allows for the identification of areas with similar patterns compared to the query area, e.g., areas with similar environmental conditions [@Jasiewicz2014LandscapeSimilarity].
Finally, the grouping of spatial patterns is a many-to-many operation in which spatial signatures are calculated for all areas and clustered based on their similarity [@Nowosad2021PatternbasedIdentification].

### 2024-10-11 Sections/03-connectivity.qmd
Starting to work on Review

Landscape connectivity describes how landscape features facilitate or impede movement [@Taylor2006LandscapeConnectivity] and can be divided into structural and functional connectivity.
It underpins ecosystem functionality, maintains biodiversity and populations, and plays an important role in many conservation actions [@CorreaAyram2016HabitatConnectivity].
Structural connectivity describes the physical arrangement of landscape elements, focusing on spatial aspects such as continuity and adjacency.
It is solely a landscape feature and independent of species characteristics [@Tischendorf2000UsageMeasurement; @CorreaAyram2016HabitatConnectivity].
Functional connectivity, in contrast, integrates landscape structure with the perceptual, behavioral, and dispersal characteristics of species, making it both species- and landscape-specific [@Adriaensen2003ApplicationLeastcost; @Taylor2006LandscapeConnectivity].

Although multiple approaches that measure connectivity have been developed [@Simpkins2018AssessingPerformance; @Keeley2021ConnectivityMetrics], the technical and conceptual quantification of connectivity has proven challenging [@Calabrese2004ComparisonshoppersGuide; @Taylor2006LandscapeConnectivity].
Generally, movement data from individuals is required to infer the elements in the landscape that individuals preferentially move through [@Simpkins2018AssessingPerformance].
While improving technologies have made tracking animal movement at high tempo-spatial resolution available [@Kays2015TerrestrialAnimal; @Nathan2022BigdataApproaches], it is still logistically challenging to track high numbers of individuals.
Thus, in practice connectivity is primarily determined by indirect estimations.
@Calabrese2004ComparisonshoppersGuide distinguished three types of estimates: i) structural connectivity determined by the physical attributes of the landscape, ii) potential connectivity as a combination of physical landscape attributes and limited information about dispersal characteristics of species, and iii) actual connectivity related to observations of individuals moving through a landscape.

@@

Many modern connectivity estimation approaches rely on resistance surfaces to represent the landscape [@Zeller2012EstimatingLandscape].
A resistance surface is a grid-based representation of a landscape where each cell is assigned a value representing the difficulty (or cost) for an individual of a particular species to traverse that cell based on landscape features such as habitat type, topography, or barriers [@Etherington2016LeastcostModelling].
Creating resistance surfaces involves obtaining landscape data for the area of interest along with a sufficiently large buffer area to avoid edge artifacts [@Dutta2022OverviewComputational] and quantifying cost values for each cell using movement data.
This can include expert opinion [widely used due to low effort, however, difficult to measure accuracy\; @Krueger2012RoleExpert], detection data (single point locations of unknown individuals), relocation data (multiple sequential locations of the same individual but at low frequency), pathway data (high-frequency relocation data allowing for movement track inference), or genetic data (samples used to calculate genetic distances between populations).

Least-cost paths assume that individuals have perfect knowledge of the landscape and will choose the path that minimizes resistance costs perfectly with no deviations [@Adriaensen2003ApplicationLeastcost].
By calculating paths between two points, it is possible to estimate connectivity between these points based on the accumulated cost along the path.
This method can also generate accumulated cost surfaces depicting the minimum cost from a single point to all other locations, helping to identify reachable areas within a threshold.
The creation of least-cost paths is a well established technique and there are highly optimized and efficient algorithms available for its calculation.

@@

Within the landscape, discrete habitat patches are modeled as nodes, and the potential movement between patches is modeled as edges.
The edges can be binary, i.e., representing whether a connection between two nodes either exists or does not exists, or edges can be weighted by cost values representing movement efforts from one patch to another [@Petsas2021LandscapeConnectivity].
Connectivity can be quantified through various graph-theoretical metrics describing different aspects, such as the importance of specific patches in maintaining connectivity.
Widely reported connectivity metrics include the probability of connectivity index [@Saura2007NewHabitat], habitat availability metrics quantifying potential and functional connectivity [@Saura2010CommonCurrency; @Pietsch2018ContributionConnectivity], or the integral index of connectivity [@Pascual-Hortal2006ComparisonDevelopment] which is based on a binary connection model [@Saura2010CommonCurrency].

@@

However, despite the robust theoretical underpinnings of graph theory and its powerful application in landscape connectivity research, contradictory results could emerge from different methods used to construct a graph or from different data sources.
In many cases, data availability limits the representation of the modelled landscape with subsequent implications for the calculated connectivity metrics and inferences of connectivity [Foltete2020CouplingLandscape].
For example, @Godet2021ComparisonLandscape modelled the same landscape using three different data types which resulted in different distribution of connectivity values.

>>>>

<!-- R1: Here connectivity seems to be described through the prism of animals, you should modify it slightly to include plants, which do not move, but disperse through different vectors, e.g. animals, gravity, wind, or water -->
<!-- R1: facilitate or impeded movement and flow -->
Landscape connectivity describes how landscape features facilitate or impede movement [@Taylor2006LandscapeConnectivity] and can be divided into structural and functional connectivity.
It underpins ecosystem functionality, maintains biodiversity and populations, and plays an important role in many conservation actions [@CorreaAyram2016HabitatConnectivity].
Structural connectivity describes the physical arrangement of landscape elements, focusing on spatial aspects such as continuity and adjacency.
It is solely a landscape feature and independent of species characteristics [@Tischendorf2000UsageMeasurement; @CorreaAyram2016HabitatConnectivity].
Functional connectivity, in contrast, integrates landscape structure with the perceptual, behavioral, and dispersal characteristics of species, making it both species- and landscape-specific [@Adriaensen2003ApplicationLeastcost; @Taylor2006LandscapeConnectivity].

Although multiple approaches that measure connectivity have been developed [@Simpkins2018AssessingPerformance; @Keeley2021ConnectivityMetrics], the technical and conceptual quantification of connectivity has proven challenging [@Calabrese2004ComparisonshoppersGuide; @Taylor2006LandscapeConnectivity].
<!-- R1: movement and dispersal data -->
<!-- R1: that individuals preferentially move [or disperse] through -->
Generally, movement data from individuals is required to infer the elements in the landscape that individuals preferentially move through [@Simpkins2018AssessingPerformance].
While improving technologies have made tracking animal movement at high tempo-spatial resolution available [@Kays2015TerrestrialAnimal; @Nathan2022BigdataApproaches], it is still logistically challenging to track high numbers of individuals.
Thus, in practice connectivity is primarily determined by indirect estimations.
@Calabrese2004ComparisonshoppersGuide distinguished three types of estimates: i) structural connectivity determined by the physical attributes of the landscape, ii) potential connectivity as a combination of physical landscape attributes and limited information about dispersal characteristics of species, and iii) actual connectivity related to observations of individuals moving through a landscape.

@@

Many modern connectivity estimation approaches rely on resistance surfaces to represent the landscape [@Zeller2012EstimatingLandscape].
<!-- R1: to traverse or disperse through that cell -->
A resistance surface is a grid-based representation of a landscape where each cell is assigned a value representing the difficulty (or cost) for an individual of a particular species to traverse that cell based on landscape features such as habitat type, topography, or barriers [@Etherington2016LeastcostModelling].
<!-- R1: Please explain what "edge artifacts" are. I could not find any explanation in the cited reference -->
Creating resistance surfaces involves obtaining landscape data for the area of interest along with a sufficiently large buffer area to avoid edge artifacts [@Dutta2022OverviewComputational] and quantifying cost values for each cell using movement data.
This can include expert opinion [widely used due to low effort, however, difficult to measure accuracy\; @Krueger2012RoleExpert], detection data (single point locations of unknown individuals), relocation data (multiple sequential locations of the same individual but at low frequency), pathway data (high-frequency relocation data allowing for movement track inference), or genetic data (samples used to calculate genetic distances between populations).

<!-- R!: Least cost path is just the result of a mathematically formulated problem, so I don't think that it assumes anything. Assumptions are rather in how such results are used in an ecological context -->
Least-cost paths assume that individuals have perfect knowledge of the landscape and will choose the path that minimizes resistance costs perfectly with no deviations [@Adriaensen2003ApplicationLeastcost].
By calculating paths between two points, it is possible to estimate connectivity between these points based on the accumulated cost along the path.
This method can also generate accumulated cost surfaces depicting the minimum cost from a single point to all other locations, helping to identify reachable areas within a threshold.
The creation of least-cost paths is a well established technique and there are highly optimized and efficient algorithms available for its calculation.

@@

Within the landscape, discrete habitat patches are modeled as nodes, and the potential movement between patches is modeled as edges.
<!-- R1: Edges can be weighted or unweighted. They also can be directed or undirected -->
The edges can be binary, i.e., representing whether a connection between two nodes either exists or does not exists, or edges can be weighted by cost values representing movement efforts from one patch to another [@Petsas2021LandscapeConnectivity].
Connectivity can be quantified through various graph-theoretical metrics describing different aspects, such as the importance of specific patches in maintaining connectivity.
Widely reported connectivity metrics include the probability of connectivity index [@Saura2007NewHabitat], habitat availability metrics quantifying potential and functional connectivity [@Saura2010CommonCurrency; @Pietsch2018ContributionConnectivity], or the integral index of connectivity [@Pascual-Hortal2006ComparisonDevelopment] which is based on a binary connection model [@Saura2010CommonCurrency].

@@

However, despite the robust theoretical underpinnings of graph theory and its powerful application in landscape connectivity research, contradictory results could emerge from different methods used to construct a graph or from different data sources.
<!-- R1: I don't think that it is specific to graphs. Any data-dependent model will be sensitive to input data. This highlights the importance of data -->
In many cases, data availability limits the representation of the modelled landscape with subsequent implications for the calculated connectivity metrics and inferences of connectivity [@Foltete2020CouplingLandscape].
For example, @Godet2021ComparisonLandscape modelled the same landscape using three different data types which resulted in different distribution of connectivity values.

### 2024-10-11 Sections/05-simulations.qmd
Starting to work on Review

Pattern-based approaches simulate landscapes without assuming any underlying abiotic or biotic processes [i.e., neutral landscape model\; @Gardner1987NeutralModels].
The earliest neutral models are based on percolation theory and assign LULC classes randomly to cells in the landscape [@Gardner1987NeutralModels], or hierarchical models that consider different spatial scales while assigning cell values [@ONeill1992HierarchicalNeutral].
Landscapes characterized by continuous values can be simulated by fractal models, such as Brownian motion [@Palmer1992CoexistenceSpecies].
Borrowing from computer graphics, more recent neutral landscape models make use of spectral synthesis [e.g., Perlin noise\; @Etherington2022PerlinNoise], or binary space partitioning [@Etherington2022BinarySpace].
Neutral landscape models are also able to simulate landscapes dominated by anthropogenic activities based on least-cost networks [@Etherington2024LeastcostNetwork].
In order to ensure realistic neutral landscapes, several approaches exist that use comparisons with real landscapes [@Inkoom2017DesigningNeutral], or are based on target values optimization [@vanStrien2016ImprovedNeutral; @Justeau-Allaire2022FragmentedLandscape].
In order to make multiple neutral landscape model algorithms accessible, several software libraries are available including Python [NLMpy\; @Etherington2015NLMpyPython], R [NLMR\; @Sciaini2018NLMRLandscapetools], Julia [NeutralLandscapes.jl\; @Poisot2023NeutralLandscapesJulia], or stand-alone software [GradientLand\; @Cambui2015GradientLandSoftware].

@@

The digital twins concept include three connected elements, namely a physical object or system, its digital representation, and a data exchange between the former two [@Jones2020CharacterisingDigital; @deKoning2023DigitalTwins], and there is an increasing interest and use of digital twins in both the industry and academia [@Jones2020CharacterisingDigital].
The need for timely and evidence-based decision making in combination with increasing data availability make digital twins a powerful tool for landscape ecology [@deKoning2023DigitalTwins].
Digital twins are increasingly applied in agricultural landscapes research, e.g., for livestock farming, controlled environment farming, or fertilization management [@Cesco2023SmartAgriculture; @Purcell2023DigitalTwins].
Further applications include planning of rural ecological landscapes [@Tan2024DigitalTwin], or exploration of relationships between urban expansion and vegetation coverage [@Zhao2022ApplyingDigital].
The use of general data processing, statistical and mechanistic models, or artificial intelligence allows digital twins to constantly update the digital representation and analyze, visualize, or predict the physical counterpart.
Thus, digital twins aim to provide real-time state of nature measurements, early detection of conservation trends, relationships of ecosystem trends and environmental conditions, impact assessments of interventions, or to identify uncertainties and information gaps [@deKoning2023DigitalTwins].

>>>>

Pattern-based approaches simulate landscapes without assuming any underlying abiotic or biotic processes [i.e., neutral landscape model\; @Gardner1987NeutralModels].
The earliest neutral models are based on percolation theory and assign LULC classes randomly to cells in the landscape [@Gardner1987NeutralModels], or hierarchical models that consider different spatial scales while assigning cell values [@ONeill1992HierarchicalNeutral].
Landscapes characterized by continuous values can be simulated by fractal models, such as Brownian motion [@Palmer1992CoexistenceSpecies].
Borrowing from computer graphics, more recent neutral landscape models make use of spectral synthesis [e.g., Perlin noise\; @Etherington2022PerlinNoise], or binary space partitioning [@Etherington2022BinarySpace].
Neutral landscape models are also able to simulate landscapes dominated by anthropogenic activities based on least-cost networks [@Etherington2024LeastcostNetwork].
<!-- R1: Note that there, although the methods in reference 146 effectively rely on target values optimization (i.e. the model aims to get as close as possible to target values), the methods in reference 147 are rather based on target value satisfaction (the model guarantees that target values are strictly satisfied, or guarantees that they cannot be satisfied) -->
In order to ensure realistic neutral landscapes, several approaches exist that use comparisons with real landscapes [@Inkoom2017DesigningNeutral], or are based on target values optimization [@vanStrien2016ImprovedNeutral; @Justeau-Allaire2022FragmentedLandscape].
In order to make multiple neutral landscape model algorithms accessible, several software libraries are available including Python [NLMpy\; @Etherington2015NLMpyPython], R [NLMR\; @Sciaini2018NLMRLandscapetools], Julia [NeutralLandscapes.jl\; @Poisot2023NeutralLandscapesJulia], or stand-alone software [GradientLand\; @Cambui2015GradientLandSoftware].

@@

The digital twins concept include three connected elements, namely a physical object or system, its digital representation, and a data exchange between the former two [@Jones2020CharacterisingDigital; @deKoning2023DigitalTwins], and there is an increasing interest and use of digital twins in both the industry and academia [@Jones2020CharacterisingDigital].
The need for timely and evidence-based decision making in combination with increasing data availability make digital twins a powerful tool for landscape ecology [@deKoning2023DigitalTwins].
Digital twins are increasingly applied in agricultural landscapes research, e.g., for livestock farming, controlled environment farming, or fertilization management [@Cesco2023SmartAgriculture; @Purcell2023DigitalTwins].
Further applications include planning of rural ecological landscapes [@Tan2024DigitalTwin], or exploration of relationships between urban expansion and vegetation coverage [@Zhao2022ApplyingDigital].
<!-- R1: You could also mention the increasing availability and accessibility of sensors deployed by drones along with data processing tools (e.g. drone photogrammetry, LiDAR) which enable high-frequency, high-resolution monitoring of 2D and 3D landscape models -->
The use of general data processing, statistical and mechanistic models, or artificial intelligence allows digital twins to constantly update the digital representation and analyze, visualize, or predict the physical counterpart.
Thus, digital twins aim to provide real-time state of nature measurements, early detection of conservation trends, relationships of ecosystem trends and environmental conditions, impact assessments of interventions, or to identify uncertainties and information gaps [@deKoning2023DigitalTwins].
